\section{Architectures}
Software architectures show the logical organizazion of a distributed system into software components, the system architecture is the final instantiation of the theoretical software architecture. \\
A couple of well known architectural styles are:
\begin{itemize}
    \item Layered
    \item Object oriented
    \item Resource orienteted
    \item Event based
\end{itemize}
Now we'll rapidly go through each of them. \\
\textbf{Layered architecture} \\
Components are organized in a layered fashion, a component can make a downcall (as in the case of calls to the operating system). \\
Upcalls are exceptions.
\smallSpace
\textbf{Object oriented architecture} \\
The components in an object oriented architecture are objects and connect to each other through procedure calls, the calls can be distributed if the objects are on different machines. \\
In the distributed case we can have separation of interfaces and objects and we can do calls via RPC. \\
Service oriented architectures are distributed applications constructed via service composition. \\
\smallSpace
\textbf{Resource base architectures} \\
A REST architecture is focused on resources and their indexing. Usually we have an interface that exposes a series of methods concerning actions we can do with respect to them. In a REST architecture:
\begin{itemize}
    \item Resources are identified through a single naming scheme
    \item All services offer the same interface
    \item Messages sent to or from service are fully self-described
    \item The communication is stateless
    \item The operations we can do with resources are a few and are standard.
\end{itemize}
When building a software architecture we are in front of two very different structures. Centralized organization, in which we have a server offering and a client recieving, the communication is of type request reply. \\
An example of centralized organization is a multitiered architecture, in which we have a client and one or more servers handling different aspects of the service (as an example: Application server, Database server). \\
